### 2022-04-23 greta/mcmc.Rmd
generalizing functions a bit

```{r}
library(tidyverse)
set.seed(1234567)

np.clip <- function(x, a, b) {
  if(x < a) return(a)
  if(x > b) return(b)
  x
}

# View the bistable curve: 
# curve( (0.75 - x)**2 * (0.25 - x), 0, 1)

step <- function(N, eta,  a = 0.75, b = 0.25) {
  N <- N + (a - N)**2 * (b - N) + eta
  N <- np.clip(N, 0, 1)
  N
}

# simulate
simulate <- function(N_init = 0.75,
                     t_max = 250L,
                     mu =0, # -5e-4,
                     sigma = 1e-2,
                     a = 0.75, 
                     b = 0.25) {
  eta <- rnorm(t_max, mu, sigma)
  N   <- numeric(t_max)
  N[1] <- N_init
  for (t in 1:(t_max-1)) {
    N[t+1] <- step(N[t], eta[t], a = a, b = b)
  }
  tibble::tibble(t = 1:t_max, N = N)
}

sims <- purrr::map_dfr(1:10, \(i) simulate(), .id = "i")
sims |> ggplot(aes(t, N, group=i)) + geom_line(alpha=0.06)
```

```{r}
gsims <- sims |> group_by(i) |> mutate(xt1 = lead(N)) |> filter(t<max(t))

x_t <- gsims$N
x_t1 <- gsims$xt1

library(greta)
a <- uniform(0, 10)
b <- uniform(0, 10)
sigma <- uniform(0, 10)
mean <-  x_t + (a - x_t) ^ 2 * (b - x_t)
distribution(x_t1) <- normal(mean, sigma )
m <- model(a, b, sigma)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("mcmc_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 10000, warmup = 5000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
```

```{r}
posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(100)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate() ,.id = "i")

bind_rows(sims, posterior_sims, .id = "model")|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.3)

@@

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- posterior_sims |> pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> select(-t) |> as.matrix()

# Score just the 1st replicate "true" observations
observed <- sims |> filter(i == 1) |> pull(N)
scores(observed, dat)

```
```{r}
# score over all replicates:

rep_scores <- sims |> group_by(i) |> group_map(~ scores(.x$N, dat)) |> bind_rows()
```

>>>>

```{r}
library(tidyverse)
library(greta)
library(bayesplot)
source("R/saddle_node.R")
```

```{r}
set.seed(4242)
pars = list(r = 1,
            K = 1,
            s = 0.1,
            h_init = .15,
            alpha = 0.0015,
            mu = 0,
            sigma = 0.02,
            t_init = 0
          )
t_max <- train_t_max + test_t_max

simulate <- simulate_sn
```

```{r}
train_reps <- 100
train_t_max <- 75
test_t_max <- 75
test_reps <- 100

sims <- purrr::map_dfr(1:train_reps, 
                       \(i) simulate(t_max=t_max, pars = pars), 
                       .id = "i")
train <- sims |> filter(t <= train_t_max)
test <- sims |> filter(t > train_t_max)
```

```{r}
m <- greta_model_sn(train)

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 100000, warmup = 80000, chains = 4, verbose = FALSE)
})
```

```{r}
bayesplot::mcmc_trace(draws)
plot_posteriors(draws, pars)
```

```{r}
fc <- forecast(draws, train, test, simulate)

fc |> ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.2)

@@

```{r}

# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()
})
```
